### 2021-10-13 plos_latex_template.tex
Rocío: shortening financial rule

Conference budgets are limited and rely mostly on sponsorship and attendance fees. Some expenses are more or less fixed but allocation of resources has to be intentional to support the goals of the conference. It is important to estimate the costs for these inclusive practices and define your priorities in advance. This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}). It also includes additional support for attendees: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team. Don't be afraid to innovate and resist the institutional inertia. "We have never payed for this, and this has always been like that" will not take the conference towards structural change.
%ast should we talk about the pandemic? something like: 
With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.

Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012} and other early career stages, such as postdoc or trainee. The definition of such positions and the associated work conditions vary for each country. Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere. In addition to this, low rates in strong currencies can represent significant amounts of money in other currencies. You can adjust the registration fee by the cost of living in each country using conversion factors (e.g., from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example). In general, people should be allowed to locate themselves in a category they consider affordable, even with the possibility of a "pay what you can" approach. Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}. Offering fee waivers to part of the participants is also a good option, but even then, other costs can be prohibitive.

Offering scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and some conferences ask for cover letters or applications to them. These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses. For online conferences, granting fee waivers is easier, and diversity scholarships can be replaced with broader participation support.
However, these applications may be counterproductive if they attract people that do not need the money but apply to enrich their CVs, and filter out people who do not feel entitled to earn them. Applying for loans, grants, and scholarships is an emotionally demanding task. Simplify applications as much as you can and be specific about the target audience of these programs. The broader the support, the simpler the application.

>>>>

% Rocío: main idea: Budgets are limited and need to choose priorities.
Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
While some expenses might be more or less fixed, % Rocío: can we give examples?
allocation of resources has to be intentional to support the goals of the conference towards inclusion. 
It is important to estimate the costs for these inclusive practices and define your priorities in advance, like paying the organizing team, necessary training (e.g. code of conduct), and accessibility practices (e.g. captions). 
Aditional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). 
Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
Instead of depending on registration fees, ask sponsors to support specific items in your inclusion expenses.

% Rocío: main idea: help take the burden out of participants for more inclusive participation
Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
(see https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example of conversion rates based on country of origin and career status). 
In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}. 
Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
Scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and conferences usually ask for cover letters or applications from them. 
These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses. 
For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper. 
Lower costs for attendance and offering financial support instead of prestigious grants--that could look good in a CV--may reduce the number of people applying for them and allow you to help all applicants. 
Applying for loans, grants, and scholarships might be an emotionally demanding task, so if this scenario is likely, simplify the process of asking for financial support.

% Rocío: Let's keep this in mind for the 10th rule or conclusion
% While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team. Don't be afraid to innovate and resist the institutional inertia. "We have never payed for this, and this has always been like that" will not take the conference towards structural change.
% %ast should we talk about the pandemic? something like: 
% With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
% Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.
